### 2017-04-05 research.Rmd

---
title: "Jeff Goldsmith"
---

<link rel="stylesheet" href="styles.css" type="text/css" />

## Motion kinematics

(text placeholder)

## Wearable devices

(text placeholder)