### 2017-04-08 teaching.Rmd
added link from teaching page to linear regression site

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />

<br>

## Current courses

Fall 2017: Data Science I

## Past courses

Spring 2016: [Linear Regression Models](http://jeffgoldsmith.com/P8111/)  
Spring 2015: Linear Regression Models  
Summer 2015: Applied Regression II  
Spring 2014: Linear Regression Models  
Spring 2014: Categorical Data Analysis  
Spring 2013: Linear Regression Models  
Spring 2013: Categorical Data Analysis

## Short courses

2017: ["Variable Selection Methods in Functional Regression"](http://jeffgoldsmith.com/IWAFDA/) (at the International Workshop on Advances in Functional Data Analysis)  
2014: "Functional Data Analysis: Techniques and Applications" (at ENAR)