### 2017-04-10 best_pubs.Rmd
adding "best pubs" page

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<br>

This page serves as a placeholder for a "best publications" page, with content to be populated shortly.