### 2017-04-10 best_pubs.Rmd
added complete citations

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<br>

Below is a collection of papers that best represents my work in statistical methods development and transdisciplinary research. For each, there is a brief synopsis and a link to download a PDF of the paper. A comprehensive list of all my publications is [here](papers.html).
<br><br>

#### Generalized Multilevel Function-on-Scalar Regression and Principal Component Analysis

In the study that motivates this paper, participants wore an accelerometer to record physical activity data continuously for five days. I treat these data as exponential-family response curves, and develop methods for function-on-scalar regression and functional principal components analysis.

**J. Goldsmith**, V. Zipunnikov, J. Schrack (2015).
Generalized Multilevel Functional-on-Scalar Regression and Principal Component Analysis.
<i> Biometrics, </i>
<b>71</b> 344-353.
[[pdf](BestPubs/GenMFPCA.pdf)].

#### Variable Selection in the Functional Linear Concurrent Model

This paper is also motivated by wearable device data. Participants wore an ambulatory blood pressure monitor that recorded observations roughy every half-hour over the course of one day. Accompanying each measurement are survey responses regarding mood, location, behavior, activity, and other variables. I posed a functional linear concurrent model for these data, and developed variable selection techniques to estimate model parameters.

**J. Goldsmith**, J. E. Schwartz.
Variable Selection in the Functional Linear Concurrent Model (2017+).
<i> Statistics in Medicine, </i>
accepted.
[[pdf](BestPubs/VBVS.pdf)].

#### Assessing Systematic Effects of Stroke on Motor Control using Hierarchical Function-on-Scalar Regression

Stroke patients were asked to make repeated reaches toward eight targets in a two-dimensional planar motion task, and their fingertip position was recorded during the execution of each reach. For these data I developed a Bayesian heirarchical function-on-scalar regression model, implemented a variational approximation for fast model building, and used posterior sampling for inference.

**J. Goldsmith**, T. Kitago (2016).
Assessing Systematic Effects of Stroke on Motor Control using Hierarchical Function-on-Scalar Regression.
<i> Journal of the Royal Statistical Society: Series C, </i>
<b>65</b> 215-236.
[[pdf](BestPubs/BayesPFSR.pdf)].

#### Modeling motor learning using heteroskedastic functional principal components analysis

A related experiment used kinematic data to quantify motor learning in healthy subjects. Working with a student, I proposed a novel extension of functional principal components analysis in which score variances depend on covariates and subject random effects. A second revision of this paper will be submitted in the coming weeks.

D. Backenroth, **J. Goldsmith**, M. D. Harran, J. C. Cortes, J. W. Krakauer, and T. Kitago.
Modeling motor learning using heteroskedastic functional principal components analysis.
<i> Journal of the American Statistical Association, </i> acceptable after major revisions.
[[pdf](BestPubs/MotorLearning.pdf)].

#### Corrected Confidence Bands for Functional Data Using Principal Components

This paper addresses an issue for inference in functional principal components analysis. Briefly, usual approaches to inference treat estimated principal components as fixed; I demonstrated that the variability in these estimates can substantially affect the coverage of confidence bands derived as if the principal components were known. I also propose a bootstrap-based procedure for obtaining bands with nominal coverage.

**J. Goldsmith**, S. Greven, C. M. Crainiceanu (2013).
Corrected Confidence Bands for Functional Data Using Principal Components.
<i> Biometrics, </i>
<b>69</b> 41-51.
[[pdf](BestPubs/FPCA_Bands.pdf)].

#### New Insights into Activity Patterns in Children, Found Using Functional Data Analyses

I develop statistical methods in response to challenges in real data, and work to disseminate those methods to a broad community. In this paper, I described function-on-scalar regression for researchers in physical activity measurement and quantification, and apply this approach to a study of activity patterns in childrean. The paper was accompanied by robust, user-friendly software, and was published in the leader subject-area journal.

**J. Goldsmith**, X. Liu, J. S. Jacobson, A. Rundle (2016).
New insights into activity patterns in children, found using functional data analyses.
<i> Medicine and Science in Sports and Exercise, </i>
<b>48</b> 1723-1729.
[[pdf](BestPubs/HeadStart.pdf)].