### 2017-04-20 git.Rmd
embedded slides for content

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<br>

This page has some stuff about Git / GitHub / GitHub with R, and is the basis for a computing club meeting.

People will get the most out of that meeting if they've installed relevant software (git, RStudio, a git client) and created a GitHub account before showing up. Instructions for doing all that can be found in Chapters 4-9 of [this book](http://happygitwithr.com) (note -- all the rest of those chapters are also awesome).

## Some slides

<script async class="speakerdeck-embed" data-id="b1f65532f0644b8fb5301f8fd6934b01" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/git-slides" title="Git stuff" target="_blank">Git stuff</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div>

After going through these slides we'll shift to an interactive demostration.

## Other materials

There's a lot of stuff to help with git in general, and with git for R in particular. This is a non-exhaustive list:

* Jenny Bryan's [book](http://happygitwithr.com) (same as the link above)
* The [GitHub](https://guides.github.com/activities/hello-world/) guide
* Version control with [RStudio](https://support.rstudio.com/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN)