### 2019-06-09 software.Rmd
add tidyfun to software

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<link rel="stylesheet" href="academicons/css/academicons.min.css"/>

<br>

## `refund`

I'm lead author of the `refund` package ([CRAN](https://cran.r-project.org/web/packages/refund/index.html); [GitHub](https://github.com/refunders/refund)). I've made contributions to several functions in this package:

* `pfr()` for penalized scalar-on-function regression
* `fpca.sc()` for functional principal components analysis
* `ccb.fpc()` for improved inference for functional principal components analysis
* `bayes_fosr()` for Bayesian function-on-scalar regression

## `tidyfun`

Together with Fabian Schiepl, I am an author of the [`tidyfun`](https://fabian-s.github.io/tidyfun/) package. This package seeks to provide accessible and well-documented software that **makes functional data analysis in `R` easy** – specifically data wrangling and exploratory analysis.

The `tidyfun` webpage includes several illustrative vignettes, and [this short overview](tidyfun.html) is a quick place to get started.

## `refund.shiny`

I worked with Julia Wrobel on the `refund.shiny` package ([CRAN](https://cran.r-project.org/web/packages/refund.shiny/index.html); [GitHub](https://github.com/refunders/refund.shiny); [paper](http://onlinelibrary.wiley.com/doi/10.1002/sta4.109/abstract)).

Through the `plot_shiny()` function, this package produces interactive graphics for several of the most common functional data analyses:

* Functional principal components analysis
* Multilevel and time-varying functional principal components analysis
* Function-on-scalar regression
* Functional linear concurrent regression

This package is closely aligned with the `refund` package: analyses are conducted using functions in `refund` and return objects that can be plotted using `refund.shiny::plot_shiny()`.

## GitHub packages

Code for several projects has been developed on [GitHub](https://github.com/jeff-goldsmith):

* This website
* [`vbvs.concurrent`](https://github.com/jeff-goldsmith/vbvs.concurrent), a package for fitting the functional linear concurrent regression model
* [`gfpca`](https://github.com/jeff-goldsmith/gfpca), a package for generalized functional principal components analysis
* [Materials](https://github.com/jeff-goldsmith/gfpca) for a short course on variable selection in functional regression models