### 2019-06-09 tidyfun.Rmd
tidyfun example! mostly for a conference

---
title: "`tidyfun`"
output:
  html_document: 
    toc: true
    toc_float: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

This page hosts a presentation / quick intro to `tidyfun` -- enjoy!

## Some slides

<script async class="speakerdeck-embed" data-id="f0d309fcbf2c4d5d9d121eee4a6144ee" data-ratio="1.77777777777778" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/tidyfun" title="tidyfun" target="_blank">tidyfun</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>

## Example usage

The code in this page is drawn from [`tidyfun`](https://fabian-s.github.io/tidyfun/) vignettes, and is intended as a quick introduction; for more details, please read the complete documentation!

If you haven't installed `tidyfun` the code below will do so.

```{r, eval = FALSE}
devtools::install_github("fabian-s/tidyfun")
```

Next I'll load the package, as well as the `tidyverse`.

```{r}
library(tidyfun)
library(tidyverse)
```

### Illustrative datasets

The datasets used in this vignette are the `tidyfun::chf_df` and `tidyfun::dti_df` dataset. The first contains minute-by-minute observations of log activity counts (stored as a `tfd` vector called `activity`) over seven days for each of 47 subjects with congestive heart failure. In addition to `id` and `activity`, we observe several covariates.

```{r view_chf}
data(chf_df)
```

A quick plot of these data is below.

```{r plot_chf}
chf_df %>% 
  ggplot(aes(y = activity)) + 
    geom_spaghetti(alpha = .05)
```

The `tidyfun::dti_df` contains ractional anisotropy (FA) tract profiles for the corpus callosum (cca) and the right corticospinal tract (rcst), along with several covariates.

```{r view_dti}
data(dti_df)
```

A quick plot of the `cca` tract profiles is below.

```{r plot_dti}
dti_df %>% 
  ggplot(aes(y = cca)) + 
    geom_spaghetti(alpha = .05)
```

### `tf` vectors

**`tf`** is a new data type for (vectors of) functional data:

- an abstract superclass for functional data in 2 forms:
    - as (argument, value)-tuples: subclass **`tfd`**, also irregular or sparse
    - or in basis representation: subclass **`tfb`**

- Includes additional attributes that define *function-like* behavior:
    - how to **evaluate** the given "functions" for new arguments
    - their **domain** 
    - the **resolution** of the argument values

First I'll pull a `tf` vector from the `tidyfun::dti_df` dataset, and then I'll extract a simple 5-element vector of functions interpolated to be on a regular grid. The resulting vector contain fractional anisotropy tract profiles for the corpus callosum (`cca`). When printed, `tf` vectors show the first few `arg` and `value` pairs for each subject.

```{r load-dti}
data("dti_df")

cca_five = dti_df$cca[1:5]
cca_five = tfd(cca_five, arg = seq(0, 1, l = 93), interpolate = TRUE)

cca_five
```

For illustration, we plot the vector `cca_five` below.

```{r, ex-fig}
plot(cca_five)
```

Functional data in basis representation:

- represented as a list of **`coefficients`** and a common **`basis_matrix`** of basis function evaluations on a vector of `arg`-values.
- contains a **`basis`** function that defines how to compute the basis for new **`arg`**s and how to differentiate or integrate it.
- (internal) flavors: 
    - `tfb_spline`: uses `mgcv`-spline bases 
    - `tfb_fpc`: uses functional principal components

It's possible to convert between types.

```{r}
cca_five_b = 
  cca_five %>% 
  tfb()
```

A quick plot illustrates the difference visually; lots of things are different in the background!

```{r}
plot(cca_five)
lines(cca_five_b, col = "red")
```

### Operations on `tf` vectors

The following is a brief overview of the kinds of operations available for `tf` vectors.

You can perform basic arithmetic and logical comparisons:

```{r}
cca_five[1] + cca_five[1] == 2 * cca_five[1]
log(exp(cca_five[2])) == cca_five[2]
cca_five - (2:-2) != cca_five 
```

You can summarize using `mean`, `sd`, and other functions:

```{r}
c(mean = mean(cca_five), sd = sd(cca_five))
```

You can smooth data using `tf_smooth`; this differs from converting to subclass `tfb`:

```{r}
cca_five_smooth = 
  cca_five %>% 
  tf_smooth("lowess")

plot(cca_five)
lines(cca_five_smooth, col = "blue")
```

You can determine whether a function satisfies a logical condition anywhere:

```{r}
cca_five %>% 
  tf_anywhere(value > .65)
```

And you can zoom in on regions of interest:

```{r}
cca_five_zoom = 
  cca_five %>% 
  tf_zoom(.5, 1)

plot(cca_five_zoom)
```

### `tf` vectors in dataframes

The main goal of `tidyfun` is to ease exploratory analysis by putting functional data in data frames. Since `tf` vectors are treated the same way as vectors of class `numeric` or `factor`, they can enter dataframes the same way.

The DTI data, for example, include scalar covariates and two functional variables:

```{r}
dti_df
```

And the CHF data is an example of a multilevel dataset with a functional observation:

```{r}
chf_df 
```

Now our functional data are tidy! That is, we have data rectangles, and each functional observation exists in a single "cell".

### Data wrangling

Dataframes using `tidyfun` to store functional observations can be manipulated using tools from `dplyr`, including `select` and `filter`:

```{r}
chf_df %>% 
  select(id, day, activity) %>% 
  filter(day == "Monday") %>% 
  ggplot(aes(y = activity)) + 
  geom_spaghetti(alpha = .05)
```

Operations using `group_by` and `summarize` are allowed:

```{r}
chf_df %>% 
  group_by(day) %>% 
  summarize(mean_act = mean(activity)) %>% 
  ggplot(aes(y = mean_act, color = day)) + 
  geom_spaghetti()
```

One can `mutate` functional observations -- here we exponentiate the log activity counts to obtain original recordings:

```{r}
chf_df %>% 
  mutate(exp_act = exp(activity)) %>% 
  ggplot(aes(y = exp_act)) + 
  geom_spaghetti(alpha = .05)
```

Some `dplyr` functions are most useful in conjunction with new functions in `tidyfun`. For example, one might use `filter` with `tf_anywhere` to filter based on the values of observed functions:

```{r}
chf_df %>% 
  filter(tf_anywhere(activity, value > 9)) %>% 
  ggplot(aes(y = activity)) + 
  geom_spaghetti()
```

One can add smoothed versions of existing observations using `mutate` and `tf_smooth`:

```{r}
chf_df %>% 
  filter(id == 1) %>% 
  mutate(smooth_act = tf_smooth(activity)) %>% 
  ggplot(aes(y = smooth_act)) + 
  geom_spaghetti()
```

One can also extract observations over a subset of the full domain using `mutate` and `tf_zoom`:

```{r}
chf_df %>% 
  filter(id == 1) %>% 
  mutate(daytime_act = tf_zoom(activity, 360, 1200)) %>% 
  ggplot(aes(y = daytime_act)) + 
  geom_spaghetti()
```

In general, EDA for functional data using `tidyverse` tools is now possible, and is often most powerful when paired with new functions in `tidyfun`.

### Visualization

We've seen both `plot` and `geom_spaghetti` to aid in understanding some content to this point -- base R for `tf` vectors, `ggplot` for tidy data. You can use more advanced options and combine with data wrangling steps.

You can use facetting:

```{r}
chf_df %>% 
  filter(day %in% c("Monday", "Sunday")) %>% 
  ggplot(aes(y = activity, color = gender)) + 
  geom_spaghetti(alpha = .1) + 
  facet_grid(~ day)
```

Together with data manipulation tools, the integration with `ggplot` can produce useful exploratory analyses. Note that this plot also introduces `geom_meatballs()`:

```{r}
chf_df %>% 
  group_by(gender, day) %>% 
  summarize(mean_act = mean(activity)) %>% 
  mutate(smooth_mean = tfb(mean_act)) %>% 
  filter(day %in% c("Monday", "Sunday")) %>% 
  ggplot(aes(y = smooth_mean, color = gender)) + 
  geom_spaghetti(size = 1.25, alpha = 1) + 
  geom_meatballs(aes(y = mean_act), alpha = .1) + 
  facet_grid(~ day)
```

Lasagna plots are a variant on a heatmaps which show functional observations in rows and use color to illustrate values taken at different arguments. In `tidyfun`, lasagna plots are implemented through `gglasagna`. A first example, using the CHF data, is below.

```{r}
chf_df %>% 
  filter(day %in% c("Monday", "Sunday")) %>% 
  gglasagna(activity)
```

A somewhat more involved example, demonstrating the `order` argument and taking advantage of facetting, is next.

```{r, dti-fig2}  
dti_df %>% 
  gglasagna(
    y = cca, 
    order = tf_integrate(cca, definite = TRUE), 
    arg = seq(0,1, l = 101)) + 
  theme(axis.text.y = element_text(size = 6)) + 
  facet_wrap(~ case, ncol = 2, scales = "free")
```

### Conversion

The `DTI` data in the `refund` package has been a popular example in functional data analysis. In the code below, we create a data frame (or `tibble`) containing scalar covariates, and then add columns for the `cca` and `rcst` track profiles. This code was used to create the `tidyfun::dti_df` dataset included in the package.

```{r}
dti_df = tibble(
  id = refund::DTI$ID, 
  sex = refund::DTI$sex, 
  case = factor(ifelse(refund::DTI$case, "MS", "control")))

dti_df$cca = tfd(refund::DTI$cca, arg = seq(0,1, l = 93))
dti_df$rcst = tfd(refund::DTI$rcst, arg = seq(0, 1, l = 55))
```

"Long" format data frames containing functional data include columns containing a subject identifier, the functional argument, and the value each subject's function takes at each argument. There are also often (but not always) non-functional covariates that are repeated within a subject. For data in this form, we use `tf_nest` to produce a data frame containing a single row for each subject.

An example is the pig weight data from the `SemiPar` package, which is a nice example from longitudinal data analysis. This includes columns for `id.num`, `num.weeks`, and `weight` -- which correspond to the subject, argument, and value.

```{r}
data("pig.weights", package = "SemiPar")

pig.weights = as_tibble(pig.weights)

pig.weights
```

We create `pig_df` by nesting weight within subject. The result is a data frame containing a single row for each pig, and columns for `id.num` and the `weight` function.

```{r}
pig_df = 
  pig.weights %>% 
  tf_nest(weight, .id = id.num, .arg = num.weeks)

pig_df
```

Additional functions allow conversion from other data structures to `tf` vectors, and also implement conversions back to these data structures.

## Next steps

`tidyfun` itself is a work in progress:

* There are several open issues / to dos / bug fixes
* We have some ideas for new features
* User feedback will help!

Integration with analysis is longer-term goal

* Some `refund` functions have been unofficially updated to work with dataframe / `tf` intpus
* More robust / complete integration is needed