### 2019-11-25 teaching.Rmd
link to this workshop kinda thing

---
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<link rel="stylesheet" href="academicons/css/academicons.min.css"/>

<br>

## Current courses

Fall 2019: [Data Science I](http://p8105.com)

## Past courses

Fall 2018: [Data Science I](http://p8105.com)  
Fall 2017: [Data Science I](http://p8105.com)  
Spring 2016: [Linear Regression Models](http://jeffgoldsmith.com/P8111/)  
Spring 2015: Linear Regression Models  
Summer 2015: Applied Regression II  
Spring 2014: Linear Regression Models  
Spring 2014: Categorical Data Analysis  
Spring 2013: Linear Regression Models  
Spring 2013: Categorical Data Analysis

## Short courses

2017: ["Functional Data Analysis for Wearables: Methods and Applications"](http://jeffgoldsmith.com/ICAMPAM/) (at the International Conference on Ambulatory Monitoring of Physical Activity)  
2017: ["Variable Selection Methods in Functional Regression"](http://jeffgoldsmith.com/IWAFDA/) (at the International Workshop on Advances in Functional Data Analysis)  
2014: "Functional Data Analysis: Techniques and Applications" (at ENAR)

## Other materials

2019: [Workshop on website development](http://jeffgoldsmith.com/ehs_workshop/index.html)
2018: Webinar: ["Data Science and Predictive Models in Public Health"]() and accompanying [reproducible lasso example](resources/lasso_example.zip) (webinar requires registration of free account)