### 2025-02-26 index.Rmd
update bio etc

---
title: "Jeff Goldsmith"
---

<link rel="stylesheet" href="styles.css" type="text/css">
<link rel="stylesheet" href="academicons/css/academicons.min.css"/>
<link rel="apple-touch-icon" href="/images/JG_transparent.png" >

<br><img src="images/jeff.jpg" alt="jeff" style="width:65%; border:10px solid"><br>

#### About

Dr. Jeff Goldsmith is Associate Professor of Biostatistics at the Columbia University Mailman School of Public Health. He serves as the Associate Dean for Data Science at MSPH, and as the Associate Director for Education at the Columbia University Data Science Institute. His statistical work focuses on functional data analysis specifically and high-dimensional, structured data more broadly. He has developed new statistical methods for regression, dimension reduction, and unsupervised learning. Dr. Goldsmith has a broad methodological and collaborative research portfolio, encompassing neuroscience and neuroimaging, physical activity behaviors, environmental health, and cardiovascular health. He has published in a range of outlets, including Biostatistics, Biometrics, Journal of the American Statistical Association, eLife, and Neurorehabilitation and Neural Repair; and he is coauthor of the book Functional Data Analysis with R.

Dr. Goldsmith is a Fellow of the American Statistical Association and was selected as an inaugural recipient of the COPSS Emerging Leader Award. He received the Mailman School's Dean's Excellence in Leadership Award the Columbia University Presidential Award for Outstanding Teaching.

He works to incorporate data science techniques for transparency and reproducibility into biostatistical analyses. Research projects are accompanied by robust, publicly available software and analytical pipelines that ensure the reproducibility of the results. This approach is informed by his work in developing a graduate course in [data science](https://p8105.com).

In 2013, Dr. Goldsmith founded (with [Dr. Todd Ogden](https://www.mailman.columbia.edu/people/our-faculty/to166) and [Dr. Phil Reiss](https://works.bepress.com/phil_reiss/)) the Functional Data Analysis Working Group ([FDAWG](https://www.mailman.columbia.edu/research/functional-data-analysis-working-group)) at Columbia Biostat.

Dr. Goldsmith received a BS in mathematics from [Dickinson College](http://www.dickinson.edu) in 2007 and a PhD in Biostatistics from [Johns Hopkins University](http://www.jhsph.edu/departments/biostatistics/) in 2012.

<br>

#### Contact Information

Jeff Goldsmith  
[Department of Biostatistics](http://www.mailman.columbia.edu/academic-departments/biostatistics)  
[Columbia Mailman School of Public Health](http://www.mailman.columbia.edu/)  
722 West 168th Street, Room 617  
New York, New York 10032